### 2020-07-07 article_figures_FLORIAN_RICOUR.Rmd
update Markdowns

---
title: "Article - Figures"
author: "Ricour Florian"
date: "June 24 2020"
output: 
  github_document:
    toc : yes
    #number_sections: yes
    fig_width : 8
    fig_height : 6
    #keep_tex : yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.height=4)
library(dplyr)
library(plyr)
library(plotly)
library(grid)
library(gridExtra)
library(lubridate)
library(ggplot2)
library(chron)
library(ncdf4)
library(reshape2)
library(gsw)
library(suncalc)
library(latex2exp)
```

## Load a bunch of data previously computed

```{r data, echo=T}
load(file='DataPaper.RData')
```

## Chla correction with Raw + Roesler + FDOM + NPQ

```{r chla_correction_functions, include = F, results=F}

#4 Function to reinitialize dataframes after cleaning (row numbers, new IDs)
clean_remove <- function(dataframe, vector_filter){
  dataframe <- dataframe[dataframe$juld %in% vector_filter,]
  rownames(dataframe) <- NULL
  dataframe <- transform(dataframe, id=as.numeric(factor(juld)))
  dataframe <- dataframe[order(dataframe$id),]
}

########
# DATA #
########
filename <- "6903240_Mprof2020.nc"
data <- ldply(as.list(filename),function(file){

  #Opening the file in a open-only mode
  ncfile   <<- nc_open(file, write = FALSE, verbose = F, suppress_dimvals = FALSE)

  #print(file)

  #Dimensions
  N_PROF   <- ncol(ncvar_get(ncfile,"PRES"))
  N_LEVELS <- nrow(ncvar_get(ncfile,"PRES"))
  juld     <- ncvar_get(ncfile,"JULD")
  pres     <- ncvar_get(ncfile,"PRES")
  lon      <- ncvar_get(ncfile,"LONGITUDE")
  lat      <- ncvar_get(ncfile,"LATITUDE")

  FloatInfo <- list(N_PROF=N_PROF,
                    N_LEVELS=N_LEVELS,
                    juld=juld,
                    pres=pres,
                    lon=lon,
                    lat=lat)

  chladf <- ExtractVar("CHLA", FloatInfo)
  bbpdf <- ExtractVar("BBP700", FloatInfo)

  if (file == "7900591_Mprof.nc" | file == "7900592_Mprof.nc"){ 
    cdomdf <- as.data.frame(rep(NA, length(chladf$value)))
    colnames(cdomdf) <- "value"
  }else{
    cdomdf <- ExtractVar("CDOM", FloatInfo)
  }

  # for a dimension mismatch issue occurring with FLOAT WMO 6900807 (between BBP and CHLA)
  if(file == "6900807_Mprof.nc"){
    chladf <- join(bbpdf, chladf, by = c("alevel", "aprofile"))
    chladf <- chladf[,11:18]
    cdomdf <- join(bbpdf, cdomdf, by = c("alevel", "aprofile"))
    cdomdf <- cdomdf[,11:18]
  }

  id <- ncvar_get(ncfile, "PLATFORM_NUMBER")

  day_night <- DayOrNight(chladf)

  data.frame(depth    = chladf$depth,
             juld     = chladf$juld,
             fluo     = chladf$value,# fluo = chla converted from fluo data
             bbp      = bbpdf$value,
             cdom     = cdomdf$value,
             qc_fluo  = chladf$qc,
             qc_bbp   = bbpdf$qc, #no qc_cdom because qc_cdom is always 0 so we cannot exploit it
             day      = month.day.year(chladf$juld,c(1,1,1950))$day,
             month    = month.day.year(chladf$juld,c(1,1,1950))$month,
             year     = month.day.year(chladf$juld,c(1,1,1950))$year,
             DOY      = as.integer(strftime(as.Date(chladf$juld,origin = '1950-01-01'), format ="%j")),#Day Of Year
             lon      = chladf$lon,
             lat      = chladf$lat,
             dir      = chladf$dir,
             day_night = day_night$day_night,
             date_long = day_night$date,
             platform = as.numeric(unique(id))
  )
})
```

```{r chla_correction_filter, include = F, results=F}

#REMOVE BAD DATA
data <- data[-(which(data$qc_fluo == 4)),] 
rownames(data) <- NULL
data <- data[-(which(data$depth < 0)),]

#Day of deployment
data <- data[data$DOY == 88 & data$year == 2018,]

smoothed_fluo <- ldply(as.list(unique(data$juld)), function(i){
  tmp <- data[data$juld == i,]
  tmp <- mmed(tmp$fluo, 5)
  data.frame(smoothed_fluo = tmp)
})

smoothed_cdom <- ldply(as.list(unique(data$juld)), function(i){
  tmp <- data[data$juld == i,]
  tmp <- mmed(tmp$cdom, 5)
  data.frame(smoothed_cdom = tmp)
})

#REPLACE FLUO WITH SMOOTHED FLUO & CDOM
data[,3] <- smoothed_fluo
data[,4] <- smoothed_cdom
data <- clean_remove(data, unique(data$juld))
data_info <- ddply(data,~juld,summarize,
                    qc_fluo = qc_fluo[which.max(fluo)],
                    depthmax = depth[which.max(fluo)],
                    maxvalue = fluo[which.max(fluo)],
                    depthmin = depth[which.max(fluo):length(fluo)][which.min(fluo[which.max(fluo):length(fluo)])],
                    integration = sum(fluo),
                    bottomdepth = max(depth),
                    min_depth = min(depth),
                    dir = dir[1],
                    lon=mean(lon),
                    lat=mean(lat),
                    platform = platform[1],
                    day = day[1], month = month[1],
                    year = year[1], DOY = DOY[1])
rm(ncfile, smoothed_cdom, smoothed_fluo)
```

```{r chla_correction_density, include = F, results=F}

# compute density profiles
density_profiles <- ldply(as.list(filename),function(file){

  ncfile   <<- nc_open(file, write = FALSE, verbose = F, suppress_dimvals = FALSE)
  N_PROF   <- ncol(ncvar_get(ncfile,"PRES"))
  N_LEVELS <- nrow(ncvar_get(ncfile,"PRES"))
  juld     <- ncvar_get(ncfile,"JULD")
  pres     <- ncvar_get(ncfile,"PRES")
  lon      <- ncvar_get(ncfile,"LONGITUDE")
  lat      <- ncvar_get(ncfile,"LATITUDE")
  FloatInfo<-list(N_PROF=N_PROF,
                  N_LEVELS=N_LEVELS,
                  juld = juld,
                  pres=pres,
                  lon=lon,
                  lat=lat)

  #NO ADJUSTED VALUE FOR PSAL AND TEMP FOR EACH FLOAT
  tempdf <- ExtractVar("TEMP",FloatInfo)
  tempdf$value[which(tempdf$qc == 4 | tempdf$qc == 3)] <- NA
  psaldf <- ExtractVar("PSAL",FloatInfo)
  psaldf$value[which(psaldf$qc == 4 | psaldf$qc == 3)] <- NA

  #DENSITY ANOMALY BASED ON TEOS-10
  psal <- gsw_SA_from_SP(psaldf$value,psaldf$depth,psaldf$lon,psaldf$lat)
  temp <- gsw_CT_from_t(psal,tempdf$value,tempdf$depth)
  sigma <- gsw_sigma0(psal,temp)

  id <- ncvar_get(ncfile, "PLATFORM_NUMBER")

  data.frame(sigma = sigma,
             depth = tempdf$depth, 
             juld  = tempdf$juld, 
             dir   = tempdf$dir,
             lon   = tempdf$lon,
             lat   = tempdf$lat,
             month = month.day.year(tempdf$juld,c(1,1,1950))$month,
             year  = month.day.year(tempdf$juld,c(1,1,1950))$year,
             DOY   = as.integer(strftime(as.Date(tempdf$juld,origin = '1950-01-01'), 
                                         format ="%j")),
             platform = as.numeric(unique(id)))
})
density_profiles <- clean_remove(density_profiles, unique(data$juld))
data_info <- transform(data_info,id=as.numeric(factor(juld)))
density_profiles <- transform(density_profiles,id=as.numeric(factor(juld)))

#MLD computation
sigma_criteria <- 0.03
depth_ref <- 10
MLDdf <- ldply(as.list(1:length(unique(density_profiles$id))), function(i){

  tmp <- density_profiles[density_profiles$id == i,]
  rownames(tmp) <- NULL

  if(all(is.na(tmp$sigma)) == TRUE){
    MLD <- NA
    sigma_surface <- NA
  }else if(length(tmp$sigma[!is.na(tmp$sigma)==TRUE])>=2){
    sigma_surface<-NA
    sigma_surface <- approx(tmp$depth,tmp$sigma,depth_ref)$y
  }

  if(is.na(sigma_surface) == FALSE){
    MLD <- max(tmp$depth[tmp$sigma <= (sigma_surface + sigma_criteria)], na.rm = T)
  }else{
    MLD <- NA
  }

  if(is.na(MLD) == TRUE){
    sigmaMaxMLD <- NA
  }else{
    sigmaMaxMLD <- mean(tmp$sigma[which(tmp$depth == MLD)], na.rm =T) 
  }

  show(i)

  data.frame(sigma_surface = sigma_surface, MLD = MLD, sigmaMaxMLD = sigmaMaxMLD,
             juld = tmp$juld[1], lat=tmp$lat[1], lon=tmp$lon[1], day = month.day.year(tmp$juld[1],c(1,1,1950))$day,
             month = month.day.year(tmp$juld[1],c(1,1,1950))$month,
             year = month.day.year(tmp$juld[1],c(1,1,1950))$year,
             DOY = as.integer(strftime(as.Date(tmp$juld[1],origin = '1950-01-01'), 
                                       format ="%j")),
             platform = tmp$platform[1])
})

# manual adjust because density inversion
MLDdf$MLD[3] <- 23.5
```

```{r chla_correction_roesler, include = F, results=F}

# add a column to the 'profiles' dataframe with OLD_CHLA (will save the uncorrected chla, can be useful for some plots)
data$raw_fluo <- data$fluo

# Add Roesler Correction
data$fluo <- data$fluo * 0.65
data$raw_roesler <- data$fluo

# EACH PROFILES HAS BEEN VALIDATED SO FAR ==> CALIBRATION EXERCISE STARTS NOW
# LET'S DO IT ONLY FOR THE DEEP PROFILE
# FDOM-based method (Xing et al. 2017)
# keep only the validation profile 
data <- data[data$id == 3,]
data_info <- clean_remove(data_info, unique(data$juld))
MLDdf <- clean_remove(MLDdf, unique(data$juld))
data <- transform(data,id=as.numeric(factor(juld)))

#FDOM-BASED CORRECTION AND MINIMUM-OFFSET CORRECTION
FDOM_OR_MINIMUM_OFFSET_CORRECTION <- ldply(as.list(1:length(unique(data$id))), function(i){
  tmp <- data[data$id == i,]
  MaxDepth <- data_info$bottomdepth[i]#Max depth of the profile
  TopDepth <- data_info$depthmin[i]#Apparent minimum

  if(all(is.na(tmp$cdom)) == TRUE){#no cdom sensor hence minimum-offset correction procedure
    offset <- tmp$fluo[which(tmp$depth == TopDepth)]
    fluo_cor <- tmp$fluo - offset
    fluo_cor[which(tmp$depth == TopDepth):length(tmp$depth)] <- 0
    slope_fdom <- NA
    C <- NA
  }else{#FDOM-based method
    calibrange <- tmp[which(tmp$depth == TopDepth):length(tmp$depth),]
    linearMod <- lm(fluo ~ cdom, data = calibrange)
    slope_fdom <- coef(linearMod)[[2]]
    C <- coef(linearMod)[[1]]
    fluo_cor <- tmp$fluo - slope_fdom*tmp$cdom - C
  }

  data.frame(depth = tmp$depth, fluo_cor = fluo_cor, 
             slope_fdom = rep(slope_fdom, length(fluo_cor)),
             C = rep(C, length(fluo_cor)), juld = tmp$juld[1], platform = tmp$platform[1], year = tmp$year[1])
})

#USE THE FDOM-corrected data
data[,3] <- FDOM_OR_MINIMUM_OFFSET_CORRECTION$fluo_cor
data$raw_roesler_fdom <- data$fluo
```

```{r chla_correction_quenching, include = F, results=F}

# QUENCHING CORRECTION
# NOTE : This function was given to me by Marin Cornec (PhD Student at the LOV)
quenching_correction <- function(fluo,depth,MLD) {
  if(is.na(MLD) == FALSE){
    f <- fluo[!is.na(fluo) & depth <= MLD]
    d <- depth[!is.na(fluo) & depth <= MLD]
    zMax <- d[which.max(f)]
    Max <- max(f)
    Corfluo <- fluo
    #Criteria from Schmechtig et al. 2014
    if(!is.na(MLD) & min(f[d<=zMax])<=(0.9*Max)) Corfluo[depth<=zMax] <- Max
    return(Corfluo)
  }else{
    return(fluo)
  }
}

#NPQ correction adapted to daily and nightly (DN) profiles
NPQ_correction_DN_adapted <- ldply(as.list(1:length(unique(data$id))), function(i){

  tmp <- data[data$id == i,]

  if(tmp$day_night[1] == "day"){
    correction <- quenching_correction(tmp$fluo, tmp$depth, MLDdf$MLD[i]) # NPQ correction only applies for daily profiles
  }else{correction <- tmp$fluo}#for nightly profiles

  data.frame(fluo_NPQ = correction)
})
data$fluo <- NPQ_correction_DN_adapted$fluo_NPQ # all correction here except the "positivation step" : fluo here = RAW + ROESLER + FDOM + NPQ

# REPLACE < 0 of CHLA by 0 (after all corrections)
# data$fluo[which(data$fluo < 0)] <- 0
# data$fluo_npq[which(data$fluo_npq < 0)] <- 0
```

```{r correction_fig, include = T}
#HPLC DATA FROM DEPLOYMENT
depth_hplc <- c(0,30,50,70,90,100,140,200,250,500,750,1000)
chla_hplc <- c(0.6111,0.6433,0.3090,0.0675,0.0229,0.0168,0.012,0.0032,
               0.0035, 0.0023, 0.0039, 0.0042)
hplc <- as.data.frame(cbind(depth_hplc, chla_hplc))

#trick for CDOM
df <- subset(data, select = c('depth', 'cdom'))
df <- df[!duplicated(df$depth),]

deployment <- ggplot(data, aes(x = fluo, y = depth, colour = "ROESLER + FDOM + NPQ")) + geom_path(size = 0.5) +
  scale_y_reverse() + geom_path(data = data, aes(x = raw_roesler_fdom, y = depth, colour = "ROESLER + FDOM")) +
  xlab(expression(Chlorophyll~a~(mg/m^3)))+ ylab("Depth [m]") +
  geom_hline(yintercept = data_info$depthmin[1], color = "black", lty = 2) +
  annotate("text", x = .8, y = data_info$depthmin[1]-20, label = "CHLA MINIMUM") +
  geom_path(data = data, aes(x = raw_fluo, y = depth, colour = "None (RAW)")) + 
  geom_path(data = data, aes(x = raw_roesler, y = depth, colour = "ROESLER")) + 
  geom_point(data = df, aes(x = cdom/10, y = depth), colour = "black", size = .5) + 
  geom_point(data=hplc, aes(y = depth_hplc, x = chla_hplc), colour = "red", size = 2, shape = 15)  + 
  scale_color_discrete(name = "Correction") +
  theme_bw() + theme(legend.position = "none")

zoom_deployment <- ggplot(data, aes(x = fluo, y = depth, colour = "ROESLER + FDOM + NPQ")) + geom_path(size = 0.5) +
  scale_y_reverse(limits = c(100,0)) + geom_path(data = data, aes(x = raw_roesler_fdom, y = depth, colour = "ROESLER + FDOM")) +
  xlab(expression(Chlorophyll~a~(mg/m^3)))+ ylab("Depth [m]") +
  geom_path(data = data, aes(x = raw_fluo, y = depth, colour = "None (RAW)")) + 
  geom_path(data = data, aes(x = raw_roesler, y = depth, colour = "ROESLER")) + 
  geom_point(data=hplc, aes(y = depth_hplc, x = chla_hplc), colour = "red", size = 2, shape = 15)  + 
  scale_color_discrete(name = "Correction") +
  theme_bw() + theme(axis.title.y=element_blank())

grid.arrange(deployment, zoom_deployment, nrow = 1, ncol = 2)
```

# Profile shapes + fitted curves

```{r shapes, include = T}
fit_curve <- function(i){
  tmp <- init_fit[i,]
  juld <- tmp$juld
  tmp2 <- init_profiles[init_profiles$juld == juld,]
  shape <- as.character(tmp$shape)
  tmp2$fluo <- normalize(tmp2$fluo) #
  depthindex <- which.min(tmp2$depth <= 90)
  tmp2 <- tmp2[1:depthindex,]
  tab <- data.frame(x=tmp2$depth,y=tmp2$fluo)
  z <- tab$x
  # only 4 shapes (E has not been encountered)
  if(shape == "G"){
    dat <- as.data.frame(fgauss(z, tmp$Fmax, tmp$Zmax, tmp$dz))
    colnames(dat) <- "fit"
    dat$fit <- normalize(dat$fit)
    dat <- cbind(tmp2$depth, dat)
  }else if(shape == "S"){
    dat <- as.data.frame(fsigmoid(z, tmp$Fsurf, tmp$Zdemi, tmp$s))
    colnames(dat) <- "fit"
    dat$fit <- normalize(dat$fit)
    dat <- cbind(tmp2$depth, dat)
  }else if(shape == "GS"){
    dat <- as.data.frame(fgauss_sigmoid(z, tmp$Fsurf, tmp$Zdemi, tmp$s, tmp$Fmax, tmp$Zmax, tmp$dz))
    colnames(dat) <- "fit"
    dat$fit <- normalize(dat$fit)
    dat <- cbind(tmp2$depth, dat)
  }else if(shape == "GE"){
    dat <- as.data.frame(fgauss_expo(z, tmp$Fsurf, tmp$Zdemi, tmp$Fmax, tmp$Zmax, tmp$dz))
    colnames(dat) <- "fit"
    dat$fit <- normalize(dat$fit)
    dat <- cbind(tmp2$depth, dat)
  }
  ggplot(tmp2, aes(x = fluo, y = depth)) + geom_point() + scale_y_reverse(limits = c(90,0)) +
    xlab("Normalized Chla") + ylab("Depth [m]") + geom_path(data = dat, aes(x = fit, y = tmp2$depth), colour = "red") +
    #theme(text=element_text(size=12)) 
    theme_bw()

}
```

```{r profile_examples}

flabels <- c("Gaussian (G)","Sigmoid (S)","Gaussian-Sigmoid (GS)","Gaussian-Exponential (GE)")
profID  <- c(111,7,12,195)

ex1 <- fit_curve(profID[1])+ggtitle(flabels[1]) + theme(axis.title.x=element_blank())
ex2 <- fit_curve(profID[2])+ggtitle(flabels[2]) + theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
ex3 <- fit_curve(profID[3])+ggtitle(flabels[3]) 
ex4 <- fit_curve(profID[4])+ggtitle(flabels[4]) + theme(axis.title.y=element_blank())

grid.arrange(ex1, ex2, ex3, ex4, nrow = 2, ncol = 2)

```