### 2020-07-29 article_figure_ARTHUR_CAPET.Rmd
clean and update

---
title: "DCM_investigation"
author: "MAST"
date: "February 20, 2020"
output: 
  github_document:
    toc : yes
    #number_sections: yes
    fig_width : 8
    fig_height : 6
    #keep_tex : yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.height=4)
library(dplyr)
library(plyr)
library(grid)
library(gridExtra)
library(processx)
library(lubridate)
library(plyr)
library(ggplot2)
library(latex2exp)
library(pracma)
```

```{r data}
# Load data previously computed
load(file='./DataPaper.RData')

freshstart <- T
```

# Evolution of the Black Sea Deep Chlorophyll Maximum

```{r Fig3}
# compute stats per months
tmp_fit <- filter(init_fit, year %in% c(2014:2019)) 
stats_month <- ddply(tmp_fit, ~month, summarize,
                      profiles = length(unique(juld)))

non_dcm <- filter(tmp_fit, !juld %in% unique(fit$juld))
dcm <- filter(tmp_fit, juld %in% unique(fit$juld))

dcm$shape <- gsub('GS', 'GS_true', dcm$shape) 
dcm$shape <- gsub('GE', 'GE_true', dcm$shape)
dcm[dcm$shape == "G",]$shape <- "G_true"

#put it all back in a single fit dataframe
tmp_fit <- rbind(non_dcm, dcm)

tmp_stu <- ddply(tmp_fit, ~month~shape, summarize,
                count = length(unique(juld)),
                count2 = length(unique(juld)))

# continue
for (i in c(1:12)){
  tmp_stu[tmp_stu$month == i,]$count <- tmp_stu[tmp_stu$month == i,]$count/stats_month$profiles[i]
  tmp_stu[tmp_stu$month == i,]$count2 <- stats_month$profiles[i]
}

# define colours 
myColors <- c("#13306DFF", "#6B4596FF", "#403891FF", "#042333FF", "#A65C85FF", "#EB8055FF", "#F9B641FF", "#E8FA5BFF")
names(myColors) <- c("G", "GE", "GS", "S", "O", "G_true", "GS_true", "GE_true") #https://stackoverflow.com/questions/6919025/how-to-assign-colors-to-categorical-variables-in-ggplot2-that-have-stable-mappin
colScale <- scale_fill_manual(name = "shape",values = myColors)

library(scales)
# https://askcodez.com/ggplot-remplace-count-avec-un-pourcentage-dans-geom_bar.html

mylabs<-lapply(seq(12), function(i){paste0(month.abb[i],'\n',format(ddply(tmp_fit, ~month, summarize,
                count = length(unique(juld)))$count[i]))})

ggplot(tmp_stu, aes(x = month, fill = shape, y = count)) +
  geom_bar(stat='identity') +
  theme_bw() + 
  scale_x_continuous(breaks = c(1:12), labels = mylabs,name='') +
  scale_y_continuous(breaks = c(0,0.25,0.5,0.75,1), labels = scales::percent(c(0,0.25,0.5,0.75,1)),name='') +
  scale_fill_manual(name = "Best-fitting Form",values = myColors,
                    labels=list("non-DCM : Sigmoïd",
                                "non-DCM : Gaussian-Sigmoïd",
                                "non-DCM : Gaussian-Exponential",
                                "non-DCM : Gaussian",
                                "Others",
                                "DCM : Gaussian-Sigmoïd",
                                "DCM : Gaussian-Exponential",
                                "DCM : Gaussian"))

```

# Seasonal Time-Series

```{r Processing, error=FALSE, warning=FALSE, echo=FALSE, message=FALSE, results='hide', include=FALSE, eval=TRUE}

# Prepare data for Temporal Series
if (freshstart){

tmp <- init_profiles
tmp_fit <- init_fit

## MODIF to avoid ordering errors ##
#tmp_fit$dcm_or_not <- NA
fit$dcm_or_not <- 1
tmp_fit <- merge(tmp_fit, fit[,c('juld','platform','dcm_or_not')],by=c('juld','platform'), all = TRUE)
tmp_fit$dcm_or_not[is.na(tmp_fit$dcm_or_not)]<-0
#tmp_fit$dcm_or_not[which(tmp_fit$juld %in% fit$juld)] <- "DCM"
####

# Add scalr diagnostics to the complete dataframe for vertical integration
complete_tmp <- merge(tmp,tmp_fit[,c('juld','platform','MLD','Zmax','Zdemi')], by=c('juld','platform'))

optical_depth <- c(6,6,6,7.5,7,7,7.5,9,8,6.5,6,6) #1/k = optical depth ==> we have a value for each month

}
```

```{r Functions, error=FALSE, warning=FALSE, echo=FALSE, message=FALSE, results='hide', include=FALSE, eval=TRUE}

############
# Functions to apply to profiles
############

content_z<-function(z, depth, fluo,total){
  print(paste0('my total is', as.character(total)))
  ztot <- sum(approx(x=depth, y=fluo,xout = seq(0,z), rule = 2)$y, na.rm=T)
  print(paste0('my local total for Z:',as.character(z),'is', as.character(ztot)))
  diff <- 0.9*total-ztot
  return (diff)
}

# Simple function to find the lower boundary of core X% CHL content.
depth90B<-function(depth, fluo, total,p=.9){
#  print(total)
  # return the depth over which integration should be doe to capture 95 % of the total CHL content.
  loctot<-0
  z90 <-14
  while(loctot<p*total){
      z90=z90+1
      loctot = sum(approx(x=depth, y=fluo,xout = seq(0,z90), rule = 2)$y, na.rm=T)
  }
  return(z90)
}

depth90T<-function(depth, fluo, total,p=.9){
#  print(total)
  # return the depth over which integration should be doe to capture 95 % of the total CHL content.
  loctot<-0
  z90 <-100
  while(loctot<p*total){
      z90=z90-1
      loctot = sum(approx(x=depth, y=fluo,xout = seq(z90,200), rule = 2)$y, na.rm=T)
  }
  return(z90)
}

#########
# Light #
#########

grad_att <- function(PRESr,k,PAR0){
  PAR<-PRESr*0
  kint<-(c(k[1],k)+c(k,k[length(k)]))/2
  PAR[1]<-PAR0*exp(-kint[1]*PRESr[1])
  for (i in 2:length(PAR)){
    PAR[i]<-PAR[i-1]*exp(-((kint[i]+kint[i])/2    )*(PRESr[i]-PRESr[i-1]))
  }
  return(PAR)
}

Att_2band <- function(PRES,PAR0,ks0,kl0,depthP0=0) {
  ##  "CHL_2BANDS" ##
  # CHLA specific attenuation + two bandwidths
  if (depthP0 !=0){
    PAR0  = PAR0/(parts*exp(-ks0*depthP0)+(1-parts)*exp(-kl0*depthP0)  ) 
  }
  PAR0s <- PAR0*parts
  PAR0l <- PAR0*(1-parts)
  ks    <- rep(ks0,length(PRES))
  kl    <- rep(kl0,length(PRES))
  PARs  <- grad_att(PRES,ks,PAR0s)
  PARl  <- grad_att(PRES,kl,PAR0l)
  PAR<-PARs+PARl
  return(PAR)
}

p2b_start = c(PAR0 = 1500 ,  ks0 = 0.04 , kl0=0.001) 
p2b_lower = c(PAR0 = 0    ,  ks0 = 0    , kl0=0    )
p2b_upper = c(PAR0 = 2500 ,  ks0 = 3   , kl0=10)

Out_Default <- data.frame("PAR0"=NA,
                          "ks0"=NA,
                          "kl0"=NA)
parts = 0.37
DefOut_isoPAR <-NA

## Filter function to select eligible profiles
#countcheck<-1
FilterForPAR<-function(depth,PAR){
 # countcheck<<-countcheck+1
  # if (countcheck<10){
  #   print(depth)
  #   print(PAR)
  #   print(sum(is.na(PAR))>length(PAR)/2)
  #   readline(prompt="Press [enter] to continue")
  # }
  return (sum(is.na(PAR))<length(PAR)/2)
}

## First attempt derives APR @ 1m from interpolation and look for 1/100 of this value
isoPAR <- function(depth, PAR){
  require(pracma)
  # Intended for Casted DF
  if (FilterForPAR(depth,PAR)){
    out    <- try_default({
    d <- depth[which( !is.na(depth) & PAR>0.1  )]
    P <- PAR[which( !is.na(depth) & PAR>0.1  )]
             p1 <-interp1(d,P, 1)      
      z100 <- interp1(P,d, p1/100) 
      return(z100)},
      DefOut_isoPAR)
  }else{
    out <- DefOut_isoPAR
  }
  return(out)
}

## Second option 
isoPARb <- function(depth, PAR){
  require(pracma)
  # Intended for Casted DF
  if (FilterForPAR(depth,PAR)){
    out    <- try_default({
    d <- depth[which( !is.na(depth) & PAR>0.1  )]
    P <- PAR[which( !is.na(depth) & PAR>0.1  )]

    m <<-nls(PAR ~ Att_2band(depth,PAR0,ks0,kl0),
           data = data.frame(depth=depth, PAR=PAR),
           start = p2b_start, 
           lower = p2b_lower,
           upper = p2b_upper,
           algorithm = "port", 
           control = list(maxiter= 1000)
           )
    mcoef  <<- coef(m)
    PAR0   <-mcoef[1]
    PARcon <- predict(m,data.frame(depth=seq(1,200)))    
    z100 <- interp1(PARcon,seq(1,200), PAR0/100) 
      return(z100)},
      DefOut_isoPAR)
  }else{
    out <- DefOut_isoPAR
  }
  return(out)
}

```

```{r light integrated diags, error=FALSE, warning=FALSE, echo=FALSE, message=FALSE, results='hide', include=FALSE, eval=TRUE}

# To compute light diagnostics, ie. iso 1% of surface PAR. 
## Two methods are operational. (see functions isoPAR and isoPARb)
## At the moment, the second one allows more values, as the first one is not working when no light data are available near-surface.

if (freshstart){
integrated_diags_par <- ddply(initPAR, .(juld,platform), summarize,
                          Ziso1 = isoPAR(depth = depth , PAR = par),
                          Ziso1b = isoPARb(depth = depth , PAR = par)
                          )
}

```

```{r}
# Add scalr diagnostics to the complete dataframe for vertical integration
if (freshstart){
complete_tmp2 <- merge(complete_tmp,integrated_diags_par, by=c('juld','platform'), all.x = TRUE)
}

```

```{r IntegratedDiagsdf, error=FALSE, warning=FALSE, echo=FALSE, message=FALSE, results='hide', include=FALSE, eval=TRUE}
# TOTAL CHL, integration using first linear interpolation on a regular grid, then sum.
if (freshstart){
integrated_diags <- ddply(complete_tmp2, .(juld,platform), summarize,
                          # TOTAL CHL, integration using first linear interpolation on a regular grid, then sum.
                            total_chla    = sum(approx(x=depth, y=fluo,xout = seq(0,200), rule = 2)$y, na.rm=T),
                          # chl content relative to MLD depth
                            DoMLD_chla    = sum(approx(x=depth, y=fluo,xout = seq(0,200), rule = 2)$y[round(unique(MLD)):200]),
                            UpMLD_chla    = sum(approx(x=depth, y=fluo,xout = seq(0,200), rule = 2)$y[  0:round(unique(MLD))]),
                            RaMLD_chla    = UpMLD_chla/total_chla*100,
                          # Visible part based on monthly optical depth
                            optic_depth   = optical_depth[unique(month)]*3,
                            visible_chla  = sum(approx(x=depth, y=fluo,xout = seq(0, optic_depth), rule = 2)$y, na.rm=T), 
                            visible_ratio = visible_chla/total_chla*100, 
                          # Content around DCM 
                            DCMcontent    = try_default(
                              sum(approx(x=depth, y=fluo,xout = seq(round(unique(Zmax)-10), round(unique(Zmax)+10)), rule = 2)$y, na.rm=T),
                              NA, quiet = FALSE), 
                            DCMratio      = DCMcontent/total_chla*100 ,
                          # sigma @ Zdemi 
                          # by definition, sigma @ MLD follows sig surface by a fixed amont. 
                            #sigmaMLD      = sigma[which.min(depth)],
                            ZLow           = min(depth[fluo<0.01]),
                          #  Z90           = depth90b(depth = depth , fluo = fluo,total = total_chla),
                            Z90B           = depth90B(depth = depth , fluo = fluo,total = total_chla, p=0.75),
                            Z90T           = depth90T(depth = depth , fluo = fluo,total = total_chla, p=0.75),
                            sigmalow       = try_default(approx(x=depth, y=sigma, xout=ZLow, rule=2)$y,NA, quiet = FALSE),
                            sigma90B       = try_default(approx(x=depth, y=sigma, xout=Z90B, rule=2)$y,NA, quiet = FALSE),
                            sigma90T       = try_default(approx(x=depth, y=sigma, xout=Z90T, rule=2)$y,NA, quiet = FALSE),
                            sigmaMLD       = min(sigma, na.rm = TRUE),#[which.min(depth)],  !! THIS IS NOT CORRECT !!
                            sigmaMAX       = try_default(approx(x=depth, y=sigma, xout=unique(Zmax), rule=2)$y,NA, quiet = FALSE),
                            sigmaDEMI      = try_default(approx(x=depth, y=sigma, xout=unique(Zdemi), rule=2)$y,NA, quiet = FALSE), 
                            #Ziso1         = isoPAR(depth = depth , PAR = par),
                            sigmaiso1      = try_default(approx(x=depth, y=sigma, xout=unique(Ziso1b), rule=2)$y,NA, quiet = FALSE)
                          )
}
```

```{r IntegratedDiagsMerge, error=FALSE, warning=FALSE, echo=FALSE, message=FALSE, results='hide', include=FALSE, eval=TRUE}
if (freshstart){
tmp_fitc<-merge(tmp_fit ,integrated_diags, by = c('juld','platform'))

tmp_fitc<-merge(tmp_fitc ,integrated_diags_par, by = c('juld','platform'), all.x = TRUE)

tmp_fitc   <- tmp_fitc[tmp_fit$year %in% c(2014:2019),]

tmp_my <- ddply(tmp_fitc, .(month,year), summarize,
                nprof = length(dcm_or_not),
                percentage_dcm =  sum(dcm_or_not)/length(dcm_or_not)*100
                ) 
}
```

```{r}

tmp_fitco<- tmp_fitc
tmp_fitco$dcm_or_not[tmp_fitco$shape=="O"]<-1
tmp_fitc[which((tmp_fitc$month<4)|(tmp_fitc$month>10)),c('Zmax','sigmaMAX','DCMratio')]<-NA
tmp_fitco[which((tmp_fitco$month<4)|(tmp_fitco$month>10)),c('Zmax','sigmaMAX','DCMratio','chla_dcm')]<-NA
```

# Back Scattering Validation of the DCM

```{r}
dbdf<- subset(complete_tmp2, select = c('juld','platform','depth','sigma','fluo','bbp','Zmax','month','lon'))
dbdf<-merge(dbdf, subset(tmp_fitc,select = c('juld','platform','dcm_or_not','chla_dcm','sigmaMAX','sigmaMLD','MLD')), by = c('juld','platform'))
dbdf<- subset(dbdf,dcm_or_not==1)

dbdf<-subset(dbdf, (month>2)&(month<11))
dbdf<-subset(dbdf, !is.nan(Zmax))

dbdf$fluo_norm <- dbdf$fluo/dbdf$chla_dcm
dbdf$sigma_norm <- (dbdf$sigma-dbdf$sigmaMLD)/(dbdf$sigmaMAX-dbdf$sigmaMLD+.00000001)
dbdf$depth_norm <- (dbdf$depth-dbdf$MLD)/(dbdf$Zmax-dbdf$MLD+.00000001)
#dbdf$depth_norm_layers<-factor(dbdf$depth_norm_layers,levels=rev(levels(dbdf$depth_norm_layers)))

dbdf<-ddply(dbdf,.(platform, juld),mutate,
            bbpmax= mean(bbp[depth_norm<1.2], na.rm=TRUE),
            bbp_norm1 = bbp/bbpmax,
            bbp_DCM = approx(depth,bbp,Zmax)$y,
            bbp_norm2 = bbp/bbp_DCM)

dbdf$sigma_norm_layers <- cut(dbdf$sigma_norm,seq(0.05,3.05,.05))
dbdf$depth_norm_layers <- cut(dbdf$depth_norm,seq(0.05,3.05,.05))
```

```{r}
blbl<-subset(dbdf, select=c("sigma_norm_layers","fluo_norm", "depth_norm_layers","depth_norm","sigma_norm","bbp_norm1","month",'lon'))
require(reshape2)
blbl<-melt(blbl, id.vars=.(sigma_norm_layers,depth_norm_layers,month,depth_norm,sigma_norm,lon))

blbl$monthn<-factor(blbl$month, labels = month.abb[unique(blbl$month)])

blbl$variable<-as.character(blbl$variable)
blbl$variable[which(blbl$variable=="fluo_norm")]<-"Chla - norm."
blbl$variable[which(blbl$variable=="bbp_norm1")]<-"Bbp - norm."

blbl$variable<-factor(blbl$variable, levels=c("Chla - norm.","Bbp - norm."))

bigplot1<-ggplot(subset(blbl, month %in% c(3,4,5,6)), aes(x=depth_norm, y=value))+
      geom_boxplot(aes(group=depth_norm_layers), outlier.size=0.5)+
  theme_bw()+coord_flip(xlim = c(0,2), ylim=c(0,2))+
  facet_grid(monthn~variable)+
  geom_vline(xintercept  = 1)+xlab('Depth - norm.')+ylab('')

bigplot2<-ggplot(subset(blbl, month %in% c(7,8,9,10)), aes(x=depth_norm, y=value))+
      geom_boxplot(aes(group=depth_norm_layers), outlier.size=0.5)+
  theme_bw()+coord_flip(xlim = c(0,2), ylim=c(0,2))+
  facet_grid(monthn~variable)+
  geom_vline(xintercept  = 1)+xlab('Depth - norm.')+ylab('')
```

```{r Fig9, fig.width=8, fig.height=6}
grid.arrange(bigplot1+scale_x_reverse(),
             bigplot2+scale_x_reverse(), nrow=1)
```

# Depth horizons

```{r Fig5}
require(reshape2)
MyPal<-'Dark2'

# To remove/add line in the plot below, simply remove/add a variable in the list. (e.g. add 'Zdemi',)
tmp_zmax1 <- subset(tmp_fitco, select = c('month','year','dcm_or_not','shape','MLD','Zmax','ZLow','Z90B','Z90T','Ziso1b'))

tmp_zmax1$Zmax[tmp_zmax1$dcm_or_not==0]<-NA
tmp_zmax1$Zmax[tmp_zmax1$Zmax>200]<-NA
tmp_zmax1$Zdemi[tmp_zmax1$shape %in% c("O","GE","G")]<-NA

tmp_zmax <- melt(tmp_zmax1[,-which(colnames(tmp_zmax1)=='dcm_or_not')], id.vars=c('month','year','shape'))

ggplot(tmp_zmax, aes(x=month, fill=variable, color=variable)) +
    geom_smooth(aes(y=value),stat = 'summary',
#              fun.data = median_hilow,
             fun.data = mean_cl_boot,
              fun.args = list(conf.int = 0.95), alpha=0.5)+
  geom_boxplot(aes(y=value, group=interaction(month,variable)), alpha=0.2, notch = TRUE)+
        scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12)) +
          scale_y_reverse(breaks = seq(100,0,-10), minor_breaks = NULL) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab('[m]')+theme_bw()+coord_cartesian(ylim=c(100,0))+
  #scale_y_reverse()+
    scale_color_brewer(aesthetics = c("colour", "fill"),
                       type = qual ,
                       palette=MyPal,
                       name="Depth Horizons", 
                       labels=list(TeX('$z_{MLD}$'),
                                TeX('$z_{DCM}$'),
                                TeX('$z_{low}$'),
                                TeX('$z_{50,bottom}$'),
                                TeX('$z_{50,top}$'),
                                TeX('$z_{\\mathrm{PAR}\\,1%}$')))
```

## Density Horizons

```{r Fig6}
tmp_sig1 <- subset(tmp_fitco, select = c('month','sigmaMLD','sigmaMAX','sigmalow','sigma90B','sigma90T','sigmaiso1' ,'dcm_or_not','year','shape'))

tmp_sig1$sigmaMAX[tmp_sig1$dcm_or_not==0]<-NA
tmp_sig1$sigmaMAX[tmp_sig1$Zmax>200]<-NA

tmp_sig <- melt(tmp_sig1[,-which(colnames(tmp_sig1)=='dcm_or_not')], id.vars=c('month','year','shape'))

ggplot(tmp_sig, aes(x=month, fill=variable, color=variable)) +
    geom_smooth(aes(y=value),stat = 'summary',
#              fun.data = median_hilow,
              fun.data = mean_cl_boot,
              fun.args = list(conf.int = 0.95), alpha=0.5)+
  geom_boxplot(aes(y=value, group=interaction(month,variable)), alpha=0.2, notch = TRUE)+
        scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab('[kg/m3]')+theme_bw()+coord_cartesian(ylim=c(16,10))+scale_y_reverse()+
      scale_color_brewer(aesthetics = c("colour", "fill"),type = qual ,palette=MyPal,
                       name="Density Horizons", 
                       labels=list(TeX('$\\sigma_{MLD}$'),
                                TeX('$\\sigma_{DCM}$'),
                                TeX('$\\sigma_{low}$'),
                                TeX('$\\sigma_{50,bottom}$'),
                                TeX('$\\sigma_{50,top}$'),
                                TeX('$\\sigma_{\\mathrm{PAR}\\,1%}$')))+
  geom_hline( yintercept = 14.5)

## How to maintain all data even those out of axis limits : https://stackoverflow.com/questions/25685185/limit-ggplot2-axes-without-removing-data-outside-limits-zoom

```

## Vertical distribution of the CHL content

```{r Fig8}
require(reshape2)
tmp_part <- melt(subset(tmp_fitco, select = c('month','year','RaMLD_chla','DCMratio')), id.vars=c('month','year'))
#,'visible_ratio'

ggplot(tmp_part, aes(x=month, fill=variable, color=variable)) +
    geom_smooth(aes(y=value),stat = 'summary',
#              fun.data = median_hilow,
              fun.data = mean_cl_boot,
              fun.args = list(conf.int = 0.95), alpha=0.5)+
        scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12)) +
        scale_y_continuous(breaks = seq(0,100,10), minor_breaks = NULL) +
    geom_boxplot(aes(y=value, group=interaction(month,variable)), alpha=0.2)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab('[%]')+theme_bw()+
  scale_color_brewer(name='Part of tot. Chl:',
                     labels=c('above MLD','within 5m of DCM','above optical depth'),
                     aesthetics = c("colour", "fill"),type = qual ,palette=MyPal)#+ylim(c(0,100))
```

## Concentrations

```{r Fig7}
require(reshape2)
tmp_conc1<-subset(tmp_fitco, select = c('total_chla','chla_surf','chla_dcm','month','year','dcm_or_not'))

tmp_conc1$chla_dcm[tmp_conc1$dcm_or_not==0]<-NA

tmp_conc1$total_chla<-tmp_conc1$total_chla/40

tmp_conc <- melt(tmp_conc1[,-which(colnames(tmp_conc1)=='dcm_or_not')], id.vars=c('month','year'))

  ggplot(tmp_conc, aes(x=month, fill=variable, color=variable)) +
      geom_smooth(aes(y=value),stat = 'summary',
  #              fun.data = median_hilow,
                fun.data = mean_cl_boot,
                fun.args = list(conf.int = 0.95), alpha=0.5)+
          scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12)) +
      geom_boxplot(aes(y=value, group=interaction(month,variable)), alpha=0.2)+
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ylab('[mg/m3]')+theme_bw()+coord_cartesian(ylim=c(0,3.5))+
    scale_color_brewer(aesthetics = c("colour", "fill"),type = qual,
                       palette=MyPal, 
                       name="Chla concentrations",
                       labels=c("Average","Surface","DCM"))
```

## Navarro's Ratio

```{r sigma_ratio_2, echo=TRUE}
require(scales)

MLDdf_info2 <- ddply(tmp_fitco, .(year,platform), summarize,
                maxMLD = max(MLD, na.rm = T),
                sigmaMaxMLD = max(sigmaMLD, na.rm = T))

sigma_ratio2 <- ddply(tmp_fitco, .(platform, juld), mutate,
   sigmaMaxMLD = MLDdf_info2$sigmaMaxMLD[which((MLDdf_info2$year == unique(year))
                                              & (MLDdf_info2$platform == unique(platform)))],
   ratio = sigmaMAX/sigmaMaxMLD)
```

```{r ratio_spatial2b}

ddd <- subset(sigma_ratio2,(month %in% c(3,4,5,6,7,8,9,10))&(!is.na(ratio)))
ddd$cutlon <- cut(ddd$lon,breaks = c(27,34.5,42))
ddd$bibi <- factor(interaction(ddd$month,ddd$cutlon), ordered = TRUE)

discfigb1<-ggplot(ddd, 
       aes(x = month, y = ratio, color=cutlon, fill=cutlon))+
  theme_bw()+
   geom_boxplot(aes(group= bibi),
                alpha=0.2, notch = TRUE, varwidth = FALSE, position='dodge2')+
  scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12))+
  scale_color_brewer(aesthetics = c("colour", "fill"),
                     type = qual ,palette=MyPal, name="Longitude", labels=c('West','East')) +
  ylab(TeX('$\\sigma_{DCM}/\\sigma_{MLD,max}$'))

discfigb1

```

## and what about light at DCM ?

```{r paratdcm, error=FALSE, warning=FALSE, echo=FALSE, message=FALSE, results='hide', include=FALSE, eval=TRUE}
# Attempt to derive PARfraction at the DCM depth

loc <- merge(initPAR, subset(tmp_fitc,select = c("platform","juld","Zmax","Z90T")),by=c("platform","juld"))

PARpartDCM <- function(depth, PAR , Zmax){
  require(pracma)
  #require(nls2)
  if (FilterForPAR(depth,PAR)){
    out    <- try_default({
      d <<- depth[which( !is.na(depth) & PAR>0.1  )]
      P <- PAR[which( !is.na(depth) & PAR>0.1  )]

      # ? nls2
      m <<-nls(PAR ~ Att_2band(depth,PAR0,ks0,kl0),
               data = data.frame(depth=depth, PAR=PAR),
               start = p2b_start, 
               lower = p2b_lower,
               upper = p2b_upper,
               algorithm = "port", 
               control = list(maxiter= 1000)
      )
      mcoef  <<- coef(m)
      PAR0   <-  mcoef[1]
      PARcon <- predict(m,data.frame(depth=c(5,Zmax)))  
      PARpartDCM <- PARcon[2]/PAR0

      ## Other approach ##
      #PARconB <- approx(x = depth,y=PAR,xout = Zmax)$y
      #PARpartDCM <- PARconB/PAR0
      return(PARpartDCM)},
      DefOut_isoPAR)
  }else{
    out <- DefOut_isoPAR
  }
}

integrated_diags_par2 <- ddply(loc, .(juld,platform), summarize,
                          PARpartDCM = PARpartDCM(depth = depth , PAR = par, Zmax=unique(Zmax))                          )

```

```{r PARatDCM2}
#tmp_fitc<-tmp_fitc[,-which(colnames(tmp_fitc)=="PARpartDCM")]
tmp_fitcl<-merge(tmp_fitc ,integrated_diags_par2, by = c('juld','platform'), all.x = TRUE)

discfig2<-ggplot(tmp_fitcl, aes(x=month, y=PARpartDCM*100)) +
    geom_smooth(stat = 'summary',
#              fun.data = median_hilow,
             fun.data = mean_cl_boot,
              fun.args = list(conf.int = 0.95),
              alpha=0.5,
              color=brewer_pal(palette = MyPal)(2)[2],
              fill=brewer_pal(palette = MyPal)(2)[2])+
  geom_boxplot(aes(group=month), alpha=0.2, notch = TRUE, color=brewer_pal(palette = MyPal)(2)[2],fill=brewer_pal(palette = MyPal)(2)[2])+
  theme_bw()+scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12))+
  scale_y_continuous(breaks = seq(0,5,1), minor_breaks = NULL, limit=c(0,5)) +
  ylab('[% of surface PAR at DCM]')
#print(discfig2)

```

```{r PARatDCM2BIS, error=FALSE}
#tmp_fitc<-tmp_fitc[,-which(colnames(tmp_fitc)=="PARpartDCM")]
tmp_fitcl<-merge(tmp_fitc ,integrated_diags_par2, by = c('juld','platform'), all.x = TRUE)

tmp_fitcl$cutlon <- cut(tmp_fitcl$lon,breaks = c(27,34.5,42))
tmp_fitcl$bibi         <- factor(interaction(tmp_fitcl$month,tmp_fitcl$cutlon), ordered = TRUE)

# discfigb2<-ggplot(tmp_fitcl, aes(x=cutlon, y=PARpartDCM*100, color=factor(month), fill=factor(month))) +
#   geom_boxplot(aes(group=bibi), alpha=0.2, notch = TRUE)+
#   theme_bw()+#scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12))+
#   scale_y_continuous(breaks = seq(0,5,1), minor_breaks = NULL, limit=c(0,5)) +
#   scale_color_brewer(aesthetics = c("colour", "fill"),type = qual ,palette=MyPal, name="Month") +
#   ylab('[% of surface PAR at DCM]')
discfigb2<-ggplot(tmp_fitcl, 
       aes(x = month, y = PARpartDCM*100, color=cutlon, fill=cutlon))+
  theme_bw()+
   geom_boxplot(aes(group= bibi),
                alpha=0.2, notch = TRUE, varwidth = FALSE, position='dodge2')+
  scale_x_continuous(NULL, labels = month.abb, breaks = seq(1,12))+
  scale_y_continuous(breaks = seq(0,5,1), minor_breaks = NULL, limit=c(0,5)) +
  scale_color_brewer(aesthetics = c("colour", "fill"),
                     type = qual ,palette=MyPal, name="Longitude", labels=c('West','East')) +
  ylab('[% of surface PAR at DCM]')

print(discfigb2)
```

# Potential Additional sections.

## Spatial

```{r Fig4,fig.width=8,fig.height=8}

require(ggmap)
#mm <- get_map

mycoord <- c(27,40.5,42,47)
myMap_stamen <- get_map(location=mycoord,source="stamen",maptype = "watercolor", force = TRUE)

tmp_fitco$monthn<-factor(tmp_fitco$month, labels = month.abb[unique(tmp_fitco$month)])

ms<-ggmap(myMap_stamen)
ms1<-ms+geom_point(data=tmp_fitco, aes(x=lon,y=lat, color=factor(dcm_or_not)))+
  facet_wrap(~monthn, scale="fixed")+
  theme_bw()+
  scale_color_discrete(name="", labels=c('non-DCM','DCM'))+
  theme(legend.position = 'top', legend.box = "horizontal")+
  scale_x_continuous(name='', breaks = c(27,30,33,36,39), limits = c(27,42), expand = expand_scale(add = 0))+
  scale_y_continuous(name='', breaks = seq(41,47), limits = c(40.7,47), expand = expand_scale(add = 0))

ms1
```

## Inter-annual variability.

Here we repeat all plots made above, but with single lines marking interannual variability.

```{r, include=FALSE}
require(reshape2)

tmp_zmaxC<- dcast(tmp_zmax,formula = month~variable, fun.aggregate = mean, na.rm=TRUE)
tmp_zmaxA<-melt(tmp_zmaxC, id.vars=c('month'))

#levels(tmp_zmax$variable)<-levels(tmp_zmax$variable)[-which()]
#levels(tmp_zmaxA$variable)<-levels(tmp_zmax$variable)[-7]

P_years_depth <- ggplot(subset(tmp_zmax, variable !="Zdemi"), aes(x=month, fill=variable, color=variable)) +
    geom_line(data=subset(tmp_zmaxA, variable !="Zdemi"),aes(y=value),alpha =1, size=0.5, linetype=2)+
      geom_smooth(aes(y=value, group=variable),
                  stat = 'summary',
                  fun.data = mean_cl_boot,
                  fun.args = list(conf.int = 0.95), 
                  alpha=0.2, linetype=0)+
  scale_x_continuous(name='', breaks=c(), labels="", limits=c(.5,12.5))+
  ylab('[m]')+theme_bw()+
  coord_cartesian(ylim=c(90,0), expand = FALSE)+scale_y_reverse()+
  scale_color_brewer(aesthetics = c("colour", "fill"),
                       type = qual ,
                       palette=MyPal,
                       name="Depth Horizons", 
                       labels=list(TeX('$z_{MLD}$'),
                                TeX('$z_{DCM}$'),
                                TeX('$z_{low}$'),
                                TeX('$z_{50,bottom}$'),
                                TeX('$z_{50,top}$'),
                                TeX('$z_{\\mathrm{PAR}\\,1%}$')))+
#  scale_color_brewer(aesthetics = c("colour", "fill"),type = qual ,palette=MyPal, name='')+
  facet_grid(.~year)+
  theme(legend.position = 'top', legend.box = "horizontal",panel.spacing = unit(0, "lines"))+
  guides(colour = guide_legend(nrow = 1))

P_years_depth
```

```{r, include=FALSE}

tmp_sigC<- dcast(tmp_sig,formula = month~variable, fun.aggregate = mean, na.rm=TRUE)
tmp_sigA<-melt(tmp_sigC, id.vars=c('month'))

P_years_sig <-ggplot(tmp_sig, aes(x=month, fill=variable, color=variable)) +
       geom_line(data=tmp_sigA,aes(y=value),alpha =1, size=0.5, linetype=2)+
      geom_smooth(aes(y=value, group=interaction(year, variable)),stat = 'summary',
             fun.data = mean_cl_boot,
              fun.args = list(conf.int = 0.95), 
                  alpha=0.2, linetype=0)+
  scale_x_continuous(name='', breaks=c(), labels="", limits=c(.5,12.5))+
  ylab('[kg/m3]')+theme_bw()+
  coord_cartesian(ylim=c(10,16), expand = FALSE)+
  scale_y_reverse()+
  #scale_color_brewer(aesthetics = c("colour", "fill"),type = qual ,palette=MyPal)+
  scale_color_brewer(aesthetics = c("colour", "fill"),type = qual ,palette=MyPal,
                       name="Density Horizons", 
                       labels=list(TeX('$\\sigma_{MLD}$'),
                                TeX('$\\sigma_{DCM}$'),
                                TeX('$\\sigma_{low}$'),
                                TeX('$\\sigma_{50,bottom}$'),
                                TeX('$\\sigma_{50,top}$'),
                                TeX('$\\sigma_{\\mathrm{PAR}\\,1%}$')))+
  facet_grid(.~year)+
  theme(legend.position = 'top', legend.box = "horizontal",panel.spacing = unit(0, "lines"))+
          guides(colour = guide_legend(nrow = 1))

P_years_sig

```

```{r inter_parts, include=FALSE}

tmp_partC<- dcast(tmp_part,formula = month~variable, fun.aggregate = mean, na.rm=TRUE)
tmp_partA<-melt(tmp_partC, id.vars=c('month'))

P_years_part <-ggplot(tmp_part, aes(x=month, fill=variable, color=variable)) +
       geom_line(data=tmp_partA,aes(y=value),alpha =1, size=0.5, linetype=2)+
      geom_smooth(aes(y=value, group=interaction(year, variable)),stat = 'summary',
             fun.data = mean_cl_boot,
              fun.args =  list(conf.int = 0.95), 
                  alpha=0.2, linetype=0)+
  scale_x_continuous(name='', breaks=c(), labels="", limits=c(.5,12.5))+
  ylab('[%]')+theme_bw()+coord_cartesian(ylim=c(0,100), expand = FALSE)+
  scale_color_brewer(name='Part of tot. Chl:',
                     labels=c('above MLD','within 5m of DCM','above optical depth'),
                     aesthetics = c("colour", "fill"),type = qual ,palette=MyPal)+
  facet_grid(.~year)+theme(legend.position = 'top', legend.box = "horizontal",panel.spacing = unit(0, "lines"))+ guides(colour = guide_legend(nrow = 1))

P_years_part

```

```{r inter_conc, include=FALSE}

tmp_concC<- dcast(tmp_conc,formula = month~variable, fun.aggregate = mean, na.rm=TRUE)
tmp_concA<-melt(tmp_concC, id.vars=c('month'))

P_years_conc <-ggplot(tmp_conc, aes(x=month, fill=variable, color=variable)) +
       geom_line(data=tmp_concA,aes(y=value),alpha =1, size=0.5, linetype=2)+
      geom_smooth(aes(y=value, group=interaction(year, variable)),stat = 'summary',
             fun.data = mean_cl_boot,
              fun.args = list(conf.int = 0.95), 
                  alpha=0.2, linetype=0)+
  scale_x_continuous(name='', breaks=c(), labels="", limits=c(.5,12.5))+
  ylab('[mg/m3]')+theme_bw()+coord_cartesian(ylim=c(0,3), expand = FALSE)+
  scale_color_brewer(aesthetics = c("colour", "fill"),type = qual ,palette=MyPal)+
  facet_grid(.~year)+scale_color_brewer(aesthetics = c("colour", "fill"),type = qual,
                       palette=MyPal, 
                       name="Chla concentrations",
                       labels=c("Average","Surface","DCM"))+
  facet_grid(.~year)+theme(legend.position = 'top', legend.box = "horizontal",panel.spacing = unit(0, "lines"))+ guides(colour = guide_legend(nrow = 1))

P_years_conc

```

```{r Fig11, fig.height=12, fig.width=10}
grid.arrange(P_years_depth,
             P_years_sig+theme( strip.background = element_blank(),strip.text.x = element_blank()), 
             P_years_part+theme( strip.background = element_blank(),strip.text.x = element_blank()),
             P_years_conc+theme( strip.background = element_blank(),strip.text.x = element_blank()),
             ncol=1)

```